---
title: "Data Preparation Lab"
format: 
  html:
    self-contained: true
    toc: true
    toc-location: left
    #theme: yeti
    #number-sections: true
    page-layout: full
 
editor: visual
execute: 
  echo: true
  message: false
  warning: false
---

For an introduction to R packages, see [here](https://www.datacamp.com/tutorial/r-packages-guide).

## 1. Load necessary packages

```{r}
library(readxl)
library(tidyverse)
library(labelled)
```

## 2. Read in Data

-   You need to specify your own directory path.
-   The function `read_excel` is part of the package `readxl`, which was loaded into the current working environment in the previous step
-   R is capable of reading data from most formats, including [data files created in other statistical packages](https://haven.tidyverse.org/). Whether the data was prepared using Excel (in CSV, XLSX, or TXT format), SAS, Stata, SPSS, or others, R can read and load the data.

```{r}
mysurvey <- read_excel("data/Survey_results.xlsx") 

```

check the names of the data columns (i.e., variable names)

```{r}
names(mysurvey)
```

## 3. Rename the variables

-   a). rename the variables with shorter names
-   b). create an ID variable

```{r}
names(mysurvey) <- c("belief", "gender", "age", "reliattend", "belief12","alcohol","presex","homosex","polaffi")
mysurvey <- rowid_to_column(mysurvey, "ID")
```

-   c). check your work

```{r}
names(mysurvey)
head(mysurvey)
```

## 4. Create labels for the shorter variable names

-   The new variable names were too short. We need more information. A label for each variable name containing such information will be great!

-   The function `var_label` is part of the package `labelled`, which was loaded into the current working environment in the openning step

```{r}
var_label(mysurvey) <- list(
  ID         = "Individual ID",
  belief     = "Are you a Christian?",
  gender     = "What is your sex?",
  age        = "How old are you?",
  reliattend = "How often do you go to a church?",
  belief12   = "Were you a Christian when you were 12 years old?",
  alcohol    = "Largest number of drinks in one occasion in the past 30 days",
  presex     = "How much agree with ‘premarital sex is always wrong'?",
  homosex    = "How much agree with same-sex marriage?",
  polaffi    = "Political affiliation: Rep, Demo, or Ind?"
)
```

## 5. Convert all character variables to factor variables {.smaller}

-   Below I am using the pipe operator from the `tidyverse` package.
-   **`factor`** is a special data type in R. Click the [link](https://stats.oarc.ucla.edu/r/modules/factor-variables/) to see more details.
-   for factors whose values have an inherent order (e.g., `reliattend`), I re-order them using `levels` argument
-   the factor variable `belief` has a category `Not Sure`, which I recoded into missing values
-   I created a numeric version for the factor variable `reliattend`

```{r}
mysurvey <-  mysurvey %>%
  mutate_if(is.character, as.factor) %>% 
  mutate(reliattend = factor(reliattend, levels = c("Rarely or never", "Less than monthly", "Monthly but not weekly", "Weekly or more"))) %>% 
  mutate(belief = fct_recode(belief, NULL = "Not Sure")) %>% 
  mutate(attend = as.numeric(reliattend))  # construct a numeric version

```

Check the levels of the factor variable

```{r}
levels(mysurvey$reliattend)
str(mysurvey)
```
